---
title: "groupware_project"
output: html_document
date: '2022-12-24'
---

## Loading libraries
```{r , echo=TRUE}
library(dplyr)
library(caret)
library(corrplot)
library(caretEnsemble)

```

## Loading Data
```{r , echo=TRUE}
dat <- read.csv("data/pml-training.csv")
print(dim(dat))

```

## Cleaning  and splitting data
Delete columns if more than 90% of the values are NA
```{r , echo=TRUE}
dat_na <- mutate_all(dat, list(~na_if(.,"")))
dat_na <- dat_na %>%  select_if(function(col) sum(is.na(col)) < nrow(dat)*0.9)
dat_na <- select_if(dat_na, function(col) is.numeric(col))
dat_na <- select(dat_na, -c(X, raw_timestamp_part_1, raw_timestamp_part_2))
dat_na$classe <- dat$classe

print(dim(dat_na))
set.seed(12345)
inTrain <- createDataPartition(y=dat_na$classe, p=0.7, list=FALSE)
training <- dat_na[inTrain,]
testing<- dat_na[-inTrain,]

```
I decided to only use numeric features.  
I also decided to exclude features `X`, `raw_timestamp_part_1`, and `raw_timestamp_part_2`
since they are note related to the movements.


## Training

```{r , echo=TRUE}
control <- trainControl(method="cv")
mod_rpart <- train(classe ~., data=training, method = 'rpart', trControl = control)
mod_lda <- train(classe ~., data=training, method = 'lda', trControl = control)
mod_gbm <- train(classe ~., data=training, method = 'gbm', trControl = control, verbose=FALSE)
mod_knn <- train(classe ~., data=training, method = 'knn', trControl = control)
mod_rf <- train(classe ~., data=training, method = 'rf', trControl = control, verbose=FALSE)


pred_rpart <- as.factor(predict(mod_rpart,newdata=testing))
pred_lda <- as.factor(predict(mod_lda,newdata=testing))
pred_gbm <- as.factor(predict(mod_gbm,newdata=testing))
pred_knn <- as.factor(predict(mod_knn,newdata=testing))
pred_rf <- as.factor(predict(mod_rf,newdata=testing))

ref <- as.factor(testing$classe)

print(postResample(pred_rpart, ref))
print(postResample(pred_lda, ref))
print(postResample(pred_gbm, ref))
print(postResample(pred_knn, ref))
print(postResample(pred_rf, ref))

print(confusionMatrix(pred_rf, ref))

```
Random forest method produced the most accurate model.


## Prediction on the testing set
```{r , echo=TRUE}
val_dat <- read.csv("data/pml-testing.csv")
data.frame(X=val_dat$X, prediction= predict(mod_rf, newdata=val_dat))
```

## Important variables
```{r , echo=TRUE}
varImp(mod_rf)
```
As shown above, `roll_belt`, `pitch_forearm`, and `yaw_belt` were most important.

